Enabling GPU Virtualization in Cloud Environments

### 2015-10-26 paper.tex
motivation re-revisado

\section{\uppercase{State-of-the-Art}}
\label{sec:state}
Our solutions to the defficiency exposed in the previous section is based on GPU virtualization, sharing resources in order
to attain a fair balance between supply and demand. 
While several efforst with the same goal have been made in the past, as exposed next, 
none of them is so ambitious as ours.

The work in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations that a node cannot 
use more GPUs than those hosted in the machine, and idle GPUs can not be shared with other machines. The solution presented 
by gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and makes
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor and, so
does its performance. Another similar solution is presented in gCloud~\cite{diab2013dynamic}. 
iWhile this solution is not yet integrated in a Cloud Computing Manager, its main
drawback is that the code of the applications must be modified
in order to be run in the virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. 
The work introduced in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute-intensive HPC applications.

Our proposal goes further and, in addition to 
bringing solutions for all kind of HPC applications, it is aimed
to boost flexibility in the use of GPUs.

@@

Cloud computing and storage have gained wide appeal  
during the last years thanks to the easy set up 
process and the low IT invest. The services offered in the cloud are useful for companies and individuals 
that look for temporary access to the last technology without spending a 
huge amount of money. In general, the systems provided by cloud 
companies match with the those requested thanks to its flexibility.

Two different kinds of cloud approaches are currently available:
on one hand, public facilities, such as AWS or 
Microsoft Azure that allow users to chose a pre-configured system.   
On the other hand, institutions/companies 
are able to deploy their own cloud infrastructure by a software as OpenStack.

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and operating system (OS). 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VMs that are allowed to use GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also present in the cluster configurations, where 
depending on the users' necessities, 
two approaches to introduce GPGPU computation can be identified:
the conservative option, features a lot of hosts with GPUs to 
ensure that is likely to satisfy all requests; in a
more daring approach, a reduced count of some GPU-enabled nodes is available, in the hope that a large burst of GPU requests 
 does not exhaust the cluster's resources.

The use of a virtualization technology such as {rCUDA} adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed, allowing all nodes in a cluster
not only to use them as if they were local, but also to share them 
achieving a higher utilization rate and reducing the idle periods.
Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
environment can leverage and share all GPUs installed in 
the physical nodes, removing past limitations.

Our work focuses on AWS as a public 
solution, and on OpenStack software as a private cloud environment. 
In order to achieve our goal, these cloud facilities need to be provided with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
needs to be implemented.

>>>>

\section{\uppercase{State-of-the-Art}}
\label{sec:state}
Our solutions to the defficiencies exposed in the previous section relies on GPU virtualization, sharing resources in order
to attain a fair balance between supply and demand. 
While several efforts with the same goal have been initiated in the past, as exposed next, 
none of them is as ambitious as ours.

The work in~\cite{younge2013enabling} allows the VM managed
by the Xen hypervisor to access the GPUs in a physical
node, but with this solution a node cannot 
use more GPUs than those hosted in the machine, and an idle GPU can not be shared with other machines. The solution presented 
by gVirtus~\cite{gvirtus} virtualizes GPUs and makes
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor, and so
does its performance. Similar solution is presented in gCloud~\cite{diab2013dynamic}. 
While this solution is not yet integrated in a Cloud Computing Manager, its main
drawback is that the application's code must be modified
in order to run in the virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. 
The work introduced in~\cite{jungpgpu} is more mature; however,
it is only focussed on compute-intensive HPC applications.

Our proposal goes further and, in addition to 
bringing solutions for all kind of HPC applications, aims 
to boost flexibility in the use of GPUs.

@@

Cloud computing and storage have gained wide appeal  
during the last years thanks to a relatively easy configuration  
process and a low IT invest. The services offered in the cloud are useful for companies and individuals 
that look for temporary access to state-of-the-art technology without spending a 
huge amount of money. In general, the systems provided by cloud 
companies match those requested by the clients thanks to its flexibility.

Two different types of cloud approaches are currently available:
on one hand, public facilities, such as AWS or 
Microsoft Azure, that allow users to chose a pre-configured system.   
On the other hand, institutions/companies 
are able to deploy their own cloud infrastructure using a software as OpenStack.

In both cases, the power of current cloud facilities relies on the use 
of VMs which  
emulate a physical node with its own 
CPUs, storage, network interconnection, GPU and OS. 
However, providing a VM equipped with a GPU implies
that the guest host must be equipped with at least one real GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VMs that can access the GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also present in cluster configurations, where 
depending on the users' necessities, 
two approaches to enable GPGPU computation can be identified:
the conservative option integrates many hosts with GPUs to 
ensure that it is likely to satisfy all requests; in a
more daring approach, a reduced count of some GPU-enabled nodes is configured, in the hope that a large burst of GPU requests 
 does not exhaust the cluster's resources.

The use of a virtualization technology, such as {rCUDA}, adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed, allowing all nodes in a cluster
to use them as if they were local, as well as to share them, 
achieving a higher utilization rate and reducing the idle periods.
Thanks to the ``shareability'' introduced by {rCUDA}, all VMs in the cloud 
environment can leverage and access all GPUs installed in 
the physical nodes, removing past limitations.

Our work selects AWS as a public 
solution, and OpenStack as a private cloud environment. 
In order to achieve our goal, these cloud facilities need to be integrated with the {rCUDA} software. In addition, for OpenStack, a module for managing virtualized GPUs 
needs to be implemented.
